---
title: "making genera dataframe"
output: html_notebook
---

packages that will be needed: 

```{r}
library(readr)
library(tidyverse)
```

The code chunk below will make a dataframe of containing all of the csv files with taxonomic information. Before running the code, make sure that setwd(the folder with the macroinvertebrates data)

```{r}
mydir = "NEON_count-macroinvertebrates (1)"

# list of all files with taxonomic information
neon_list <- list.files(pattern = "taxonomyProcessed*", recursive = TRUE)
# there are 136 files

# reading all csv files with taxonomic information
tbl <-
    list.files(pattern = "taxonomyProcessed*", recursive = TRUE) %>% 
    map_df(~read_csv(.))

```

Below are some questions for exploring the data and determining which variables are useful for us.

```{r}
# if there were multiple labs that were involved in the identification
tbl %>%
  summarise(n_distinct(laboratoryName)) # there are 2 distinct laboratories

# how many different people were involved in the identification?
tbl %>%
  summarise(n_distinct(identifiedBy)) # 17 different people
```

We will now pull out the relevant columns.

```{r}
genera_count <- tbl %>% 
  select(siteID, collectDate, scientificName, genus, invertebrateLifeStage, 
         individualCount, subsamplePercent, estimatedTotalCount, # count columns
         indeterminateSpecies, distinctTaxon,
         laboratoryName, identifiedBy)
```

